### 2022-12-15 Scripts/trial.Rmd
Datasets

---
title: "TidyTuesdayCoffeeRatings"
author: "Andrew Couch"
date: "7/7/2020"
output: html_document
---

Video: https://www.youtube.com/watch?v=XoCgToAuzsU&list=PLJfshcspBCYeJeO8YFT5e5HxuYOb5a_1W&index=35&t=1187s

```{r}
library(tidyverse)
coffee_ratings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2020/2020-07-07/coffee_ratings.csv')
coffee_ratings %>% glimpse()
```

```{r}
coffee_ratings %>% ggplot(aes(total_cup_points)) + geom_histogram() + scale_x_log10()
```

```{r}
#Owners can own multiple companies and companies can own multiple farms 
coffee_ratings %>%
  select(owner, company, farm_name) %>% 
  count(company, farm_name, sort = TRUE) %>% 
  drop_na()
```
```{r}
#Farms can have multiple owners 
coffee_ratings %>% count(farm_name, owner) %>% 
  drop_na() %>% 
  arrange(farm_name) %>% 
  count(farm_name, sort = TRUE)
```

```{r}
#Verify that we selected the correct features for point analysis
coffee_ratings %>% 
  select(total_cup_points, country_of_origin, aroma:cupper_points) %>% 
  mutate(row_num = row_number()) %>% 
  gather(key = "key", value = "value", -total_cup_points, -country_of_origin, -row_num) %>% 
  group_by(row_num) %>% 
  summarise(value = sum(value),
            total_cup_points = mean(total_cup_points))
```

```{r}
coffee_ratings %>% 
  filter(country_of_origin == "Ethiopia") %>% 
  select(aroma:cupper_points) %>% 
  gather() %>% 
  group_by(key) %>% 
  summarise(value = mean(value)) %>% 
  ungroup() %>% 
  mutate(key = str_replace(key, "_", " ") %>% str_to_title()) %>% 
  mutate(key = fct_reorder(key, value)) %>% 
  ggplot(aes(x = key, y = value, color = key)) + 
  geom_point(size = 5) + 
  geom_segment(aes(x = key, xend = key, y = value, yend = 0)) + 
  theme(legend.position = "none") + 
  ylab("") + 
  xlab("") + 
  coord_flip() + 
  labs(title = "Avg Flavour Profile")
```

```{r}
coffee_ratings %>% 
  filter(country_of_origin == "Ethiopia") %>% 
  select(variety) %>% 
  drop_na() %>% 
  count(variety) %>% 
  mutate(variety = fct_reorder(variety, n)) %>% 
  ggplot(aes(x = n, y = variety, fill = variety)) + 
  geom_col() + 
  ylab("") + 
  labs(title = "Bean Variety") + 
  theme(legend.position = "none")
```

```{r}
library(kableExtra)
coffee_ratings %>% 
  filter(country_of_origin == "Ethiopia") %>% 
  select(total_cup_points, species, country_of_origin, region) %>% 
  group_by(species, region) %>% 
  top_n(total_cup_points, n = 1) %>% 
  ungroup() %>% 
  mutate(region = str_trunc(region, 12, "right")) %>% 
  arrange(desc(total_cup_points)) %>% 
  kable()
```

```{r}
library(broom)
coffee_ratings %>% 
  select(country_of_origin, aroma:cupper_points) %>% 
  mutate(highlight = if_else(country_of_origin == "Ethiopia", "Highlight", "No-Highlight")) %>% 
  select(-country_of_origin) %>% 
  gather(key = "key", value = "value", -highlight) %>% 
  group_by(key) %>% 
  do(t_test = t.test(value~highlight, data = .) %>% tidy()) %>% 
  unnest(t_test) %>% 
  mutate(diffference = case_when(
    conf.low < 0 & conf.high < 0 ~ "Different",
    conf.low > 0 & conf.high > 0 ~ "Different",
    TRUE ~ "Not-Different"
  )) %>% 
  mutate(key = str_replace(key, "_", " ") %>% str_to_title()) %>% 
  mutate(key = fct_reorder(key, -estimate)) %>% 
  ggplot(aes(x = key, y = estimate, color = diffference)) + 
  geom_pointrange(aes(ymin = conf.low, ymax = conf.high)) + 
  geom_hline(yintercept = 0, linetype = "dashed") + 
  coord_flip() + 
  theme(legend.position = "none") + 
  xlab("") + 
  labs(title = "How different are the flavour profiles?")
```