### 2022-12-19 Scripts/Map_study_locations.Rmd
edits

---
title: "Study locations"
output:
  html_document: default
  pdf_document: default
---

```{r, echo=FALSE, message=FALSE, cache=FALSE,  warning=FALSE}
library(gtools) #Library that allows to sort files by numerical order
library(tidyverse)
library(sf)
#Read all files with their paths
files <- dir("../Data/Clean_data", pattern="*.rds", full.names=T)
files <- mixedsort(files)

#Generate list of lists
all_data = files %>%
map(readRDS)

#Read all file names
file_names <- dir("../Data/Clean_data", pattern="*.rds", full.names=F)
file_names <- mixedsort(file_names)

#Delete extension
file_names = str_replace(file_names, ".rds", "")
#Rename all elements of the list
names(all_data) <- file_names

#Select coordinates from the list of lists----
data <- list()

for (i in 1:length(file_names)) {

  int_list <- all_data[[file_names[i]]]$InteractionData
  data[[i]] <- bind_rows(map(int_list, function(x) x %>% #map here is the same as a lapply
                select(Longitude, Latitude)),  .id = 'Internal_id')
}

#Rename list and 
names(data) <- file_names
data = bind_rows(data,  .id = 'Study_id')

#Add manual jitter to these three points
data = data %>% 
mutate(Latitude = case_when(Study_id == "16_Manicor" ~ Latitude+0.25, TRUE ~ Latitude)) %>% mutate(Latitude = case_when(Study_id == "17_Manicor" ~ Latitude+0.25, TRUE ~ Latitude)) %>%
mutate(Latitude = case_when(Study_id == "18_Manicor" ~ Latitude+0.25, TRUE ~ Latitude))

data_coord <- st_as_sf(data, coords = c(3:4)) %>% 
st_set_crs(4326) %>% 
st_transform(3035)

data_coord$Study_id <- factor(data_coord$Study_id, levels = file_names)

# install.packages("sf")
library(sf)
# install.packages("giscoR")
library(giscoR)

# Year
year_ref <- 2016

# Get all countries and transform to the same CRS
cntries <- gisco_get_countries(year = year_ref,
        resolution = 03) %>%
        st_transform(3035)

set.seed(1)

colors <- c("#8AD69E", "#74C5CE",  "#FDE9DE",  "#45A631", "#95F7F0", "#CB53E1",
"#FE90BB", "#B187E1", "#F8A982", 
"#0075DC", "#993F00",  "#5EF1F2", 
"#005C31", "#74C5CE", "black", 
"lightblue",  "darkorange3","#8F7C00",
"#8F7C00","#8F7C00", "#9DCC00", 
"#C20088", "gold1","#FFA8BB",
"#426600","#94FFB5","#003380",  
"salmon", "palegreen","greenyellow",
"mediumaquamarine", "tomato", "orange")
#To generate more random colors 
#randomcoloR::distinctColorPalette(k = length(file_names), altCol = F, runTsne = T)

# Plot
g <- ggplot() +
geom_sf(data = cntries, fill = "grey80", color = NA) +
xlim(c(2200000, 7150000)) +
ylim(c(1380000, 5500000)) +
theme(panel.grid.major = element_line(color = gray(0.5), linetype = "dashed", 
        size = 0.5), panel.background = element_rect(fill = "aliceblue"),
panel.border = element_rect(colour = "black", fill=NA, size=1)) +
geom_sf(data = data_coord, aes(color=Study_id), size = 0.8, stroke = 0, shape = 16) +
theme(legend.position="none") + 
scale_colour_manual(name = "Study \n locations", values = colors)

library(plotly)
ggplotly(g)

```