### 2022-12-28 index.qmd
quarto

---
title: "Punk Title Here"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

```{r load_things, echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(gh) #How heavy would this library be?

releases <- gh("GET /repos/PunkUserName/PunkRepoName/releases",
   owner = "PunkUserName",
   repo = "PunkRepoName")
tag <- releases[[1]][["tag_name"]]
```

## Authors

Punk Author ^1^, ...

> (1) Does Punks care about their institutions?

**Last update:** `r Sys.Date()`

## Abstract

Tell me why I need to read this, and do it fast. Also, this document is is version `r tag`.

### How to cite this:

You can cite this dynamic document: P. Author. `r substr(Sys.Date(), 1, 4)` Punk Title. version `r tag`. DOI: [10.0000/zenodo.0000000](https://doi.org/10.0000/zenodo.0000000).

### Download in PDF:

### Source code:

You can find the source code, as well as previous releases of this repositoty at: [https://github.com/PunkUserName/PunkRepoName](https://github.com/PunkUserName/PunkRepoName)