### 2024-02-01 Manuscript/Figure_2a.Rmd
Trying to show the year of the studies on figure 2. Not sure if it contains too much info now

---
title: "Untitled"
output: pdf_document
date: "2024-02-01"
---

```{r Load libraries, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
library(dplyr) #For handling fata 
library(ggplot2) #For plotting
library(giscoR) #For plotting
library(sf) #For handling coordinates and plotting
library(patchwork)#For plotting
library(ggstar) #For plotting (cool shapes)
library(scales) #For plotting (decimals on axes)
library(lubridate) #To operate with dates
library(tidyr) #Reshape data (wide format)
library(stringr) #Fix/modify strings
library(forcats)

```

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 2}. Proportion of the major pollinator orders by habitat types and bioclimatic regions in the SafeNet database. The orders, from left to right, include Hymenoptera, Diptera, Lepidoptera and Coleoptera. The horizontal barplot on the right indicates the number of studies that were conducted on each habitat type or bioclimatic region. Note that a single study can countribute to more than one habitat or bioclimatic region. Areas with a greater number of studies are more likely to depict accurate proportions of pollinators in those systems.", fig.width=6, fig.height=8}

#Read interaction data 
data = readRDS("../Data/3_Final_data/Interactions_uncounted.rds")
#Load data (done it in earlier chunks)
#Select cols of interest
data1 = data %>% 
select(Study_id, Network_id, Authors_habitat,
  Corine_land_cover, SafeNet_habitat, Bioregion,
  Latitude, Longitude, Day, Month, Year, Plant_rank,
  Plant_status, Plant_matchtype, Plant_accepted_name,Plant_order,
  Pollinator_rank, Pollinator_status, Pollinator_matchtype,
  Pollinator_accepted_name, Pollinator_order, Interaction)

#Prepare orders for filtering
main_orders = c("Hymenoptera", "Lepidoptera", "Coleoptera", "Diptera")

#Check number of studies by habitat
habitat = data %>% 
select(Study_id, SafeNet_habitat) %>% 
distinct() %>%  
group_by(SafeNet_habitat) %>% 
summarise(Habitat_count = n()) 
#Check number of studies by bioregion
bioregion = data %>% 
select(Study_id, Bioregion) %>% 
distinct() %>%  
group_by(Bioregion) %>% 
summarise(Bioregion_count = n()) 
#Check pols by habitat
pollinator_percent = data1 %>%
group_by(SafeNet_habitat, Pollinator_order) %>%
filter(Pollinator_accepted_name!= "Apis mellifera") %>% 
filter(Pollinator_order %in% main_orders) %>% 
summarise(count = n()) %>%
group_by(SafeNet_habitat) %>%
mutate(Proportion = count/sum(count)) %>%
select(-count) 
#Check pols by bioregion
pollinator_percent_bio = data1 %>%
group_by(Bioregion, Pollinator_order) %>%
filter(Pollinator_accepted_name!= "Apis mellifera") %>% 
filter(Pollinator_order %in% main_orders) %>% 
summarise(count = n()) %>%
group_by(Bioregion) %>%
mutate(Proportion = count/sum(count)) %>%
select(-count) 
#Preparing datasets for horizontal plot
a1 = habitat %>% 
mutate(Group = "Habitat") %>% 
rename(Habitat = SafeNet_habitat) %>% 
rename(Count = Habitat_count)

b1 = bioregion %>% 
mutate(Group = "Bioregion") %>% 
rename(Habitat = Bioregion) %>% 
rename(Count = Bioregion_count)

#Bind datasets
c1 = bind_rows(a1,b1)

#Create datasets and rename cols
a2 = pollinator_percent_bio %>% 
mutate(Group = "Bioregion") %>% 
rename(Habitat = Bioregion)

b2 = pollinator_percent %>% 
mutate(Group = "Habitat") %>% 
rename(Habitat = SafeNet_habitat)
#Bind rows
c2 = bind_rows(a2, b2)

#Prepare to set ordered labels
l2 = c2 %>% 
filter(Pollinator_order == "Hymenoptera") %>% 
arrange((Proportion)) %>% 
pull(Habitat)
#Set order of labels
c2$Habitat = factor(c2$Habitat, levels = l2)

#Set the right order of labels
c2$Pollinator_order = factor(c2$Pollinator_order, levels = c("Hymenoptera", "Diptera", "Lepidoptera", "Coleoptera"))

p1 = ggplot(c2, 
aes(fill=Pollinator_order, y=Proportion, x=Habitat)) + 
geom_bar(position = position_stack(reverse = TRUE), stat="identity")+
theme_bw() +
theme(legend.position = "bottom",
      plot.margin=unit(c(0,0,0,0), "mm"),
      legend.text = element_text(size=6),
      legend.key.size = unit(2, "mm"),
      legend.spacing.y = unit(0.1, "cm"),
       legend.box.spacing = unit(0, "pt"),
      axis.title = element_text(face = "bold", size = 8),
      axis.text = element_text(size=6)) +
xlab(NULL) +
scale_y_continuous(breaks = c(0.25, 0.5,0.75), labels = c("0.25", "0.5","0.75"))+
theme(strip.text.x = element_text(size = 12, face = "bold"),
    strip.background = element_rect(
color=NA, fill=NA, size=0.5, linetype="solid")) +
scale_fill_manual(values = c("#ffd67a","skyblue", "lightpink2", "tomato3"),
                  name = NULL) +
coord_flip(expand = FALSE) +
ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")), space = "free", scales = "free_y") +
guides(fill=guide_legend(nrow=2,byrow=TRUE)) +
ggtitle("(a)") +
theme(plot.title = element_text(size = 10, face = "bold",hjust = -0.25, vjust = 1))

#Unify labels
c1$Habitat = factor(c1$Habitat, levels = l2)
#Create barplot on the side
p2 =ggplot(c1, aes(Habitat, Count)) +
geom_col(fill = "black",width = 0.9) +
theme_light() +
theme(plot.margin=unit(c(0,0,0,0), "mm"),
      axis.ticks.y = element_blank(), 
  axis.text.y = element_blank(), 
      strip.text.x = element_text(size=12),
      strip.background = element_rect(
color=NA, fill=NA, size=0.5, linetype="solid"),
  axis.title = element_text(face = "bold", size = 8),
axis.text = element_text(size=6)) +
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(),
panel.border = element_blank(),
axis.line = element_line(color = 'black')) +
scale_y_continuous(breaks = c(0, 10,20), labels = c("0", "10","20"), limits = c(0,30))+
xlab(NULL) +
ylab("Number of studies")+
theme(legend.position = "none")+
coord_flip(expand = F) +
ggforce::facet_col(~ factor(Group, levels=c("Habitat","Bioregion")), space = "free", scales = "free_y")

data2 = data %>%  select(Study_id, Year) %>% 
distinct() %>% 
group_by(Year) %>% 
summarise(Counts = n())

p3 = ggplot(data2, aes(Year, Counts)) +
geom_col(fill = "gray10", width = 0.9) +
theme_light() +
theme(plot.margin=unit(c(0,0,0,0), "mm"),
  axis.title.y = element_text(face = "bold", size = 8, vjust = -2),
  axis.title.x = element_text(face = "bold", size = 8),
  axis.text.y = element_text(size = 6),
  axis.text.x = element_text(size = 6,  angle = 90), 
  strip.text.x = element_text(size=12),
  strip.background = element_rect(
  color="black", fill=NA, size=0.5, linetype="solid")) + 
ylab("Number of studies") +
coord_cartesian(expand = F, clip = "off") + 
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(),
  panel.border = element_rect(colour = "black", fill=NA, size=1)) +
theme(panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
panel.background = element_blank(),
panel.border = element_blank(),
axis.line = element_line(color = 'black')) +
scale_y_continuous(breaks = c(0, 4,8,12), labels = c("0", "4","8", "12"), limits = c(0,12)) +
scale_x_continuous(breaks = seq(min(data2$Year), max(data2$Year), by = 2), limits = c(min(data2$Year),max(data2$Year)))+
ggtitle("(b)") +
theme(plot.title = element_text(size = 10, face = "bold",hjust = -0.25, vjust = 1))

a = plot_spacer() +  p3 + plot_spacer() + plot_layout(ncol=1, heights= c(9,10,1))
#Plot panel
p1 + p2 + a +plot_layout(widths = c(4,2, 2.5),heights= c(10,10,4), ncol=3)

```