### 2024-02-29 Manuscript/Figure_2.Rmd
progress

---
output: pdf_document
date: "2024-02-29"
---

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 3}.", fig.width=10, fig.height=10}

#Load libraries
library(ggplot2)
library(ggnewscale)
library(ggtreeExtra)
library(ggtree)
library(ggpattern)
library(patchwork)
library(cowplot)

#Read data
tree_family_interactions= readRDS("../Data/Manuscript_info/bee_phylo.RData")
family_level1 = readRDS("../Data/Manuscript_info/bee_family_info.RData")
bee_plotting = readRDS("../Data/Manuscript_info/bee_plotting.RData")

#Plot
p = ggtree(tree_family_interactions,  size=0.5,  alpha=1)+ 
geom_tippoint(aes(size= Interactions, fill = Pollinator_family), color="black", shape=21) +
geom_tiplab(linetype='dashed', linesize=.05, offset = 0.19, size=2.4, fontface=4) +
labs("log(Species)")+
scale_size(name = "Species", range = c(0, 5.5)) +
guides(fill = "none") +
theme(legend.position = "bottom")

p1 = p + 
new_scale_fill() +
geom_fruit(data=family_level1, geom=geom_bar,
      mapping=aes(y=Pollinator_genus, x=Interactions1, 
      fill = Pollinator_family1),
      pwidth=0.19, 
      orientation="y", 
      stat="identity",
      color="black",
      offset = 0.23,
      axis.params=list(axis       = "x",
                         text.size= 1.8,
                         hjust    = 1,
                         vjust      = 0.5,
                         nbreak     = 2),grid.params=list())  +
theme(legend.position = "bottom")

p2 = ggplot(data = bee_plotting, aes(reorder(Pollinator_family, -Spp_number), 
       Spp_number, fill = Group, pattern = Group)) +
geom_col_pattern(position = "dodge",color = "black", pattern_fill = "black") +
scale_pattern_manual(values = c(A = "stripe", B = "none"), labels= c(A = "SafeNet spp", B = "European spp")) +
scale_fill_manual(values = c(A = "tan2", B = "black"), labels= c(A = "SafeNet spp", B = "European spp")) + 
coord_cartesian(expand = FALSE) +
theme_bw() + 
ylab("Number of species") +
xlab("")+
ggtitle("Bee family coverage") +
theme(axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, face = "bold"),
 legend.box.margin=margin(40,40,40,40),
 legend.position = "bottom")

p2_1 = plot_grid(NULL,p2,NULL, ncol = 1, rel_heights  = c(1, 1.5, 0.5),rel_widths  = c(0.1, 2, 0.1))
plot_grid(p1,p2_1, rel_widths = c(2, 2), rel_heights  = c(4, 0.4), align = "none")

```