### 2024-03-04 Manuscript/Figure_3.Rmd
phylo plants updated

---
output: pdf_document
---

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 3}.", fig.width=8, fig.height=8}

library(ggplot2)
library(ggtree)
library(ggnewscale)
library(ggtreeExtra)
library(viridis)
#Save data and plot it in an Rmarkdown
tree_family_interactions = readRDS("../Data/Manuscript_info/plant_phylo.RData")
d1 = readRDS("../Data/Manuscript_info/plant_coverage_phylo_RData")

#Check if there are differences in labels and order
ggtree(tree_family_interactions, layout="circular", size=0.15, open.angle=5, alpha=0.25) +
geom_tippoint(aes(size= Interactions, fill = Interactions), color="black", shape=21) +
geom_tiplab(linetype='dashed', linesize=.05, offset = -5, size=1.85, fontface=2, hjust=1) +
scale_size(range = c(0,4), guide = 'none') +
scale_fill_continuous(type = "viridis", direction=-1) + 
theme(legend.position = "bottom") +
guides(fill=guide_legend(), size = guide_legend()) +
new_scale_fill() +
geom_fruit(data = d1, geom=geom_bar,
   mapping=aes(y=label, x=Percent, fill = fam_group),
   pwidth=0.1, 
   orientation="y", 
   stat="identity",
   color="black",
   offset = 0.028,
   linewidth=0.35,
   axis.params=list(
     axis = "x",
     text.size  = 1.5,
     hjust  = 1,
     vjust = 0.5,
     nbreak = 3,
     fontface = 2),
  grid.params=list())+
guides(fill = "none")+
scale_fill_manual(values = c(A = "lightgrey", B = "tan2"), guide = "none") +
geom_fruit(geom=geom_bar,
    mapping=aes(y=label, x=log(Spp_Europe +1)),
    pwidth=0.1, 
    orientation="y", 
    stat="identity",
    color="black",
    offset = 0,
    linewidth=0.35) +
theme(legend.margin=unit(1,"cm"),
      legend.spacing = unit(1, "cm"),
      plot.margin = margin(4,0,4,0), 
      legend.position = c(0.5,0.09),
      legend.direction = "horizontal")

```