### 2024-03-05 Manuscript/Figure_S1.Rmd

---
output: pdf_document
date: "2024-02-29"
---

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 3}.", fig.width=10, fig.height=4}
#Prepare figure for supplementary material
#Coverage across high taxonomic ranks of bees, syrphids and butterflies

#Load libraries
library(dplyr)
library(readr)
library(viridis)
library(ggplot2)
#--------------------------------------------------------------------------------------------#
#Start from bees----
#Load data
#Safenet interactions
data = readRDS("../Data/3_Final_data/Interactions_uncounted.rds")
#Bee-syrphid Masterlist
master_list = read_csv("../Data/Species_taxonomy/Thesaurus/Master_bees_syrphids.csv")
#Load butterfly masterlist
master_list1 <- read.delim(file = "../Data/Species_taxonomy/Thesaurus/Butterfly_taxon.txt",
                    header = T)
#Bee families
bee_fam = c("Apidae", "Megachilidae", "Halictidae", 
            "Andrenidae", "Colletidae","Melittidae")

#Species per family within SafeNet
bee_spp_safenet = data %>% 
select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
filter(Pollinator_rank == "SPECIES") %>% 
filter(Pollinator_order == "Hymenoptera") %>% 
filter(!Pollinator_accepted_name == "Apis mellifera") %>% 
filter(Pollinator_family %in% bee_fam) %>% 
group_by(Pollinator_family) %>% 
summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
mutate(Group = "SafeNet")
#Species per family within Europe
bee_spp_europe =  master_list %>% 
select(Family, Genus, Species, GenusAndSpecies) %>% 
filter(Family %in% bee_fam) %>% 
group_by(Family) %>% 
summarise(Spp_number = n_distinct(GenusAndSpecies)) %>% 
rename(Pollinator_family = Family) %>% 
mutate(Group = "Europe")

#Bind both datasets for plotting
bee_fam_coverage = bind_rows(bee_spp_safenet, bee_spp_europe)

p1 = ggplot(bee_fam_coverage, aes(reorder(Pollinator_family,-Spp_number), Spp_number, fill = Group)) +
geom_bar(stat="identity", width=.5, position = "dodge", color = "black") +
coord_cartesian(expand = FALSE) +
xlab(NULL) +
theme_bw() +
theme(axis.text.x = element_text(angle = 45, hjust =1),
      panel.border = element_rect(size=1)) +
ylab("Number of species") +
scale_fill_manual(values = c(viridis(2)))+
ggtitle("Bee family coverage")
#--------------------------------------------------------------------------------------------#
#Continue with syrphids----
#As we don't have subfamilies, let's obtained them from the master list
syrphids = master_list %>% 
filter(Family == "Syrphidae") %>% 
select(Family,Subfamily,  Genus, Species, GenusAndSpecies) %>% 
rename(Pollinator_accepted_name = GenusAndSpecies)
#Syrphid species safenet
syrphid_spp_safenet = data %>% 
select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
filter(Pollinator_rank == "SPECIES") %>% 
filter(Pollinator_family== "Syrphidae") %>% 
distinct()
#Add subfamily info
syrphid_spp_safenet = left_join(syrphid_spp_safenet, syrphids)
#Summarise spp per subfamily
syrphid_spp_safenet = syrphid_spp_safenet %>% 
group_by(Subfamily) %>% 
summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
mutate(Group = "SafeNet")
#Get spp per subfamily at european level
syrphid_spp_europe = master_list %>% 
filter(Family == "Syrphidae") %>% 
select(Family, Subfamily, Genus, Species, GenusAndSpecies) %>% 
group_by(Subfamily) %>% 
summarise(Spp_number = n_distinct(GenusAndSpecies)) %>% 
mutate(Group = "Europe")
#Bind both datasets for plotting
syrphid_fam_coverage = bind_rows(syrphid_spp_safenet, syrphid_spp_europe)
#Plot
p2 = ggplot(syrphid_fam_coverage, aes(reorder(Subfamily,-Spp_number), Spp_number, fill = Group)) +
geom_bar(stat="identity", width=.5, position = "dodge", color = "black") +
coord_cartesian(expand = FALSE) +
xlab(NULL) +
ylab(NULL) +
theme_bw() +
theme(axis.text.x = element_text(angle = 45, hjust =1),
      panel.border = element_rect(size=1)) +
scale_fill_manual(values = c(viridis(2)))+
ggtitle("Syrphid subfamily coverage")
#--------------------------------------------------------------------------------------------#
#Continue with butterflies----
#Do it at family level
#Create vector with butterfly families
butt_fam = c(unique(master_list1$family))
#Species per family within SafeNet
butterfly_spp_safenet = data %>% 
select(Pollinator_rank, Pollinator_accepted_name, Pollinator_order, Pollinator_family)%>%
filter(Pollinator_rank == "SPECIES") %>% 
filter(Pollinator_order == "Lepidoptera") %>% 
filter(Pollinator_family %in% butt_fam) %>% 
group_by(Pollinator_family) %>% 
summarise(Spp_number = n_distinct(Pollinator_accepted_name)) %>% 
mutate(Group = "SafeNet")
#Species per family within Europe
butterfly_spp_europe =  master_list1 %>% 
select(family, genus, scientificName) %>% 
group_by(family) %>% 
summarise(Spp_number = n_distinct(scientificName)) %>% 
rename(Pollinator_family = family) %>% 
mutate(Group = "Europe")
#Bind both datasets for plotting
butterfly_fam_coverage = bind_rows(butterfly_spp_safenet, butterfly_spp_europe)
#Plot
p3 = ggplot(butterfly_fam_coverage, aes(reorder(Pollinator_family,-Spp_number), Spp_number, fill = Group)) +
geom_bar(stat="identity", width=.5, position = "dodge", color = "black") +
coord_cartesian(expand = FALSE) +
xlab(NULL) +
ylab(NULL) +
theme_bw() +
theme(axis.text.x = element_text(angle = 45, hjust =1),
      panel.border = element_rect(size=1)) +
scale_fill_manual(values = c(viridis(2)))+
ggtitle("Butterfly family coverage")

#--------------------------------------------------------------------------------------------#
library(patchwork)
p1+p2+p3 + plot_layout(guides = "collect") & 
theme(plot.title = element_text(size = 12, face = "bold"),
      axis.title.y = element_text(face = "bold"))

```