### 2024-05-22 Manuscript/Figures_and_Tables/Figure_6_Updated.Rmd
edits

---
output:
  pdf_document:
header-includes:
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
---

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="95%", fig.cap= "\\textbf{Figure 6}. ", fig.width=10, fig.height=7, fig.align='center'}

# Load libraries
library(dplyr)
library(ggplot2)
library(stringr)
library(patchwork)

# Load plots generated in scripts 7_2_Nestedness and 7_2_Connectance

connectance_species = readRDS("../../Data/Manuscript_info/Figure6_connectance_species.rds")
connectance_latitude = readRDS("../../Data/Manuscript_info/Figure6_connectance_latitude.rds")
nestedness_z_scores = readRDS("../../Data/Manuscript_info/nestedness_z_scores.rds")
nestedness_latitude = readRDS("../../Data/Manuscript_info/nestedness_latitude.rds")

#Prepare first connectance plots
library(betareg)
model = betareg(Connectance ~ log_geometric_mean_spp, data = connectance_species)
p1 = ggplot(connectance_species, aes(x = log_geometric_mean_spp, y = Connectance))+
geom_point(pch = 21, fill = "azure3", size=2.5) +
geom_line(aes(y = predict(model, connectance_species)), colour= "black")+
xlab("log(Species)") +
theme_bw() +
ggtitle("(a)")

p3 = connectance_latitude %>% 
ggplot(aes(Latitude, residual_conectance)) +
geom_point(aes(fill = Bioregion, shape = Bioregion), stroke = 0.25, size=2.5) +
scale_fill_viridis_d() +
scale_colour_viridis_d() +
scale_shape_manual(values = c(
  "Alpine" = 21,  # Circle
  "Atlantic" = 22, # Triangle
  "Boreal" = 23,   # Square
  "Continental" = 24,  # Diamond
  "Mediterranean" = 25  # Cross
  )) +
geom_line(data = connectance_latitude,aes(x=Latitude, y = fit), inherit.aes = FALSE)+
geom_ribbon(data = connectance_latitude, 
    aes(x=Latitude,ymin = lwr, ymax = upr), inherit.aes = FALSE,
    fill = "grey70", alpha= 0.3) +
theme_bw()+
ylab("Residual connectance") +
ggtitle("(c)")

#Prepare now nestedness plots
# Plot
p = 0.05 #cutoff probability 95% confidence
critical_value = qnorm(p/2) #double tail probability divide by 2

nestedness_z_scores = nestedness_z_scores %>%
mutate(infra_over_represented = case_when(
z_score < -abs(critical_value) ~ "Under-represented",
between(z_score, -abs(critical_value), abs(critical_value)) ~ "No statistical difference",
    z_score > abs(critical_value) ~ "Over-represented"
  ))

p2 = nestedness_z_scores %>% 
ggplot(aes(x = z_score)) +
geom_histogram(aes(y = ..density.., fill = infra_over_represented), 
               bins = 100, alpha = 0.5, color = "black", position = "stack") +
stat_function(data = nestedness_z_scores, fun = function(x) dnorm(x, mean = mean(nestedness_z_scores$z_score), sd = sd(nestedness_z_scores$z_score)) * 1.7, 
              n = 1000, inherit.aes = FALSE, color = "gray18", size = 1) +
theme_bw() +
coord_cartesian(expand = FALSE) +
geom_vline(xintercept = -abs(critical_value), linetype = "longdash", colour = "red") +
geom_vline(xintercept = abs(critical_value), linetype = "longdash", colour = "red") +
ylab("Density") +
xlab("Z scores (NODF)")  +
xlim(-6, 9) +
ylim(0, 0.8) +
scale_fill_manual(name = "Observed against null",
                  limits = c("Under-represented", "No statistical difference", "Over-represented"),
                  labels = c("Less nested", "No difference", "More nested"),
                  values = c("Under-represented" = "coral2", 
                             "No statistical difference" = "palegreen3", 
                             "Over-represented" = "cyan3")) +
ggtitle("(b)")

# 2nd Plot
p4 = nestedness_latitude %>% 
ggplot(aes(Latitude, Normalised_nestedness)) +
geom_point(aes(fill = Bioregion, shape = Bioregion), stroke = 0.25, size=2.5) +
theme_bw() +
ylab("NODFc") +
scale_fill_viridis_d() +
scale_colour_viridis_d() +
scale_shape_manual(values = c(
  "Alpine" = 21,  # Circle
  "Atlantic" = 22, # Triangle
  "Boreal" = 23,   # Square
  "Continental" = 24,  # Diamond
  "Mediterranean" = 25  # Cross
  )) +
geom_line(data = nestedness_latitude,aes(x=Latitude, y = fit), inherit.aes = FALSE)+
geom_ribbon(data = nestedness_latitude, 
    aes(x=Latitude,ymin = lwr, ymax = upr), inherit.aes = FALSE,
    fill = "grey70", alpha= 0.3) +
theme_bw()+
ggtitle("(d)")

# Layout plots
#a = (p1 + p2) 
#b = p3 + p4 + plot_layout(guides = 'collect') & theme(legend.position = "right")
#a / b

# Layout plots
a = (p1 + p2)  & theme(legend.position = "right", legend.justification = "left")
b = p3 + p4 + plot_layout(guides = 'collect') & theme(legend.justification = "left")
a / b  & theme(axis.title =  element_text(size=16, face="bold"),
               plot.title = element_text(size=22),
               panel.border = element_rect(colour = "black",
               fill=NA, size=1),
               axis.text=element_text(size=12))

```