### 2024-08-22 Manuscript/Figures_and_Tables/Exact_location.qmd
adressing comments from colleagues

---
latex-tinytex: false
format:
  pdf:
    include-in-header: 
      text: |
        \usepackage{lscape}
        \newcommand{\blandscape}{\begin{landscape}}
        \newcommand{\elandscape}{\end{landscape}}
---

```{r Load libraries, echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE}
library(dplyr) #For handling fata 
library(ggplot2) #For plotting
library(giscoR) #For plotting
library(sf) #For handling coordinates and plotting
library(patchwork)#For plotting (binding plots)
library(ggstar) #For plotting (cool shapes)
library(scales) #For plotting (decimals on axes)
library(tidyr)
library(viridis)
```

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="110%", fig.cap= "Figure 1"}

#Load data
data = readRDS("../../Data/3_Final_data/Interaction_data.rds")

#Fix dates
data = data %>%
dplyr::mutate(Year = lubridate::year(Date), 
                Month = lubridate::month(Date), 
                Day = lubridate::day(Date))

#Prepare coordinates
for_plotting = data %>% 
select(Study_id, Latitude, Longitude, Bioregion) %>% 
distinct()
#Get countries and transfrom coordinates
countries <- gisco_get_countries(
resolution = 20) %>%
st_transform(3035) %>% 
sf::st_as_sf(crs = 4326)

#Try to plot both together (same plot)
map = ggplot(countries) +
geom_sf(fill="floralwhite",color = "black", 
        alpha = 1,linewidth=0.1) + 
geom_point(data = for_plotting, alpha=0.75,size=1.2,pch=21,stroke = 0.25,aes(x= Longitude, y=Latitude, fill= Bioregion))+  
coord_sf(crs = st_crs(3035), default_crs = st_crs(4326),
         xlim = c(-14, 39), ylim = c(37, 70))+
theme(panel.grid.major = element_blank(), 
     legend.key = element_blank(), 
     legend.text = element_text(size = 6), 
     legend.title = element_text(face ="bold",  size = 8),
     legend.spacing.x=unit(-0.5, "lines"), 
     legend.key.width=unit(0.5, "lines"), 
     legend.box.margin=margin(-10,-10,-10,-10),
     axis.text = element_text(size=6,face="bold"),
     axis.title = element_text(face="bold", size = 8),
) +
theme(panel.background = element_rect(fill = "aliceblue"),
        panel.border = element_rect(colour = "black",
        fill=NA, size=1)) + 
 scale_size_continuous(name= "Species counts",range = c(0, 3.75))  +
ylab("Latitude") +
xlab("Longitude") +
scale_x_continuous(breaks = seq(0, 20, by = 10)) +
scale_y_continuous(breaks = seq(40, 60, by = 10)) +
guides(fill = guide_legend(override.aes = list(size = 1.5, byrow = TRUE))) +
scale_fill_manual(values = c("purple3", 
                               "dodgerblue3", 
                               "cyan4", 
                               "forestgreen",
                               "orange2",
                               "tomato3",
                               "khaki2"))

#Plot everything
map 
```