### 2024-10-02 Manuscript/Figures_and_Tables/Figure_S6_NULL.Rmd
final edits

---
output:
  pdf_document:
header-includes:
- \usepackage{caption}
- \captionsetup[figure]{labelformat=empty}
- \captionsetup[table]{labelformat=empty}
---

```{r echo=FALSE, message=FALSE, cache=FALSE, warning=FALSE, out.width="90%", fig.width=12, fig.height= 8, fig.cap="\\textbf{Figure S6}."}

library(dplyr)
library(ggplot2)

#Read data
null_metrics_networks = readRDS("../../Data/Working_files/null_metrics_networks_nullmodel2.rds")
metrics_by_network = readRDS("../../Data/Working_files/metrics_by_network.rds")

#Finally, find % of networks that differ statistically
# Merge and process data
d = left_join(metrics_by_network, null_metrics_networks, by = "Network_id")
d = d %>% 
mutate(z_score = (Classic_nestedness - Mean_null_classic_nestedness) / Deviation_null_classic_nestedness)

p = 0.05 #cutoff probability 95% confidence
critical_value = qnorm(p/2) #double tail probability divide by 2

d1 = d %>%
mutate(infra_over_represented = case_when(
z_score < -abs(critical_value) ~ "Under-represented",
between(z_score, -abs(critical_value), abs(critical_value)) ~ "No statistical difference",
    z_score > abs(critical_value) ~ "Over-represented"
  ))

#Calculate percentages
z_percent = d1 %>% 
group_by(infra_over_represented) %>% 
summarise(percent = length(infra_over_represented)/nrow(.) * 100)

# Plot
 d1 %>% 
  ggplot(aes(x = z_score)) +
  geom_histogram(aes(y = ..count.., fill = infra_over_represented), 
                 bins = 50, alpha = 0.5, color = "black", position = "stack") +
  theme_bw() +
  coord_cartesian(expand = FALSE) +
  geom_vline(xintercept = -abs(critical_value), linetype = "longdash", colour = "red") +
  geom_vline(xintercept = abs(critical_value), linetype = "longdash", colour = "red") +
  ylab("Number of networks") +
  xlab("Z scores (NODFc)")  +
 # xlim(-6, 9) +
 # ylim(0, 0.8) +
  scale_fill_manual(name = "Observed against null",
                    limits = c("Under-represented", "No statistical difference", "Over-represented"),
                    labels = c("Less nested", "No difference", "More nested"),
                    values = c("Under-represented" = "coral2", 
                               "No statistical difference" = "palegreen3", 
                               "Over-represented" = "cyan3")) +
xlim(-6,6) +
theme(plot.title = element_text(size = 24),
      text = element_text(size = 20),
      axis.title  = element_text(face="bold"))

```